Demystifying the Combination of Dynamic Slicing and \\ Spectrum-based Fault Localization

### 2019-02-20 paper/ijcai/paper.tex
small fixes

Model-based software diagnosis (MBSD)~\cite{REITER198757,DEKLEER200325} and
Spectrum-Based Fault Localization (\sfl{}) are two very popular techniques that
leverage different principles to automate fault isolation. An example of a MBSD
is Dynamic Slicing (DS)~\cite{Agrawal:1990:DPS:93542.93576} as well as some
work~\cite{wotawa2002model,Mayer:2008:EMM:1642931.1642950,mayer2008prioritising,Perez:2018:LQR:3304889.3304927,Ko:2008:DRA:1368088.1368130},
a technique that traces back statements in the code that influence some point of
interest, such as the evaluation of a failing assertion. In contrast to \ds{},
\sfl{}~\cite{7390282} does not analyze code. Instead, it is a black-box,
statistical fault localization method that computes suspiciousness values
associated with program entities (e.g., methods) based on coverage information
gathered during the execution of test cases as well as the output of test cases.
More precisely, \sfl{} uses coverage information of passing and failing test
cases to identify likely faulty statements and produces on output a list of
program entities ranked in decreasing order of suspiciousness. \ds{} and \sfl{}
are complementary debugging techniques~\cite{DBLP:conf/sac/AbreuGZG08}.
Intuitively, \ds{} identifies irrelevant parts of the code whereas \sfl{} ranks
the relevant parts of the code.

@@

The first question addresses the important problem of missing faulty
statements. Naturally, the consequence of that problem for debugging
is that developers would never be able to locate faults. Different
slicers can miss faulty statements for different reasons. In this
paper, we used Critical Slicing~\cite{DeMillo:1996:CSS:229000.226310},
which, in principle, could miss statements because of imprecise
oracles that guide the slicing process. We evaluated how often that
issue manifests in our experiments. The second research questions
effectively evaluates the impact of the combination. Our
results show that the combination of the two approaches improves the accuracy
of fault localization of 51\% of the studied faults.

The contributions of this work are the following ones:
\begin{itemize}
	\item A study, using real-world applications and bugs, on the combination
    of \ds{} and \sfl{} for bug localization of \emph{Java} faulty programs.
   \item A tool that implements the combination of \ds{} with \sfl{}, coined
   \comb{}.
\end{itemize}

@@

Several types of spectra exist.  The most commonly used is called
hit-spectrum, where the coverage matrix is encoded in terms of binary
\emph{hit} (1) and \emph{not hit} (0) flags, \ie{}, $\A{ij} = 1$ if
$t_i$ covers $c_j$ and $\A{ij} = 0$ otherwise.  \sfl{} takes as input
the pair $(\mathcal{A},e)$ and produces on output a list of components
ranked by their faulty suspiciousness. To that end, the first step of
the technique consists of determining what columns of the matrix $A$
resemble the error vector $e$ the most.  For that, an intermediate
component frequency aggregator $n_{pq}(j)$ is computed $n_{pq}(j) =
|\{i\mid \A{ij}=p \wedge e_i=q\}|$. $n_{pq}(j)$ denotes the number of
runs in which the component $j$ has been active during execution ($p =
1$) or not ($p=0$), and in which the runs failed ($q = 1$) or passed
($q = 0$).  For instance, $n_{11}(j)$ counts the number of times
component $j$ has been involved ($p = 1$) in failing executions ($q =
1$), whereas $n_{10}(j)$ counts the number of times component $j$ has
been involved in passing executions. We then calculate similarity to
the error vector by means of applying \emph{fault predictors} to each
component to produce a score quantifying how likely it is to be
faulty.  Components are then ranked according to such likelihood
scores and reported to the user. Ochiai is one of those fault
predictors that has shown to perform
well~\cite{7390282,Pearson:2017:EIF:3097368.3097441}. The Ochiai
formula is given by the following equation
$\textit{ochiai}\,$=$\,\N{11}/\sqrt{(\N{11}+\N{01}) * (\N{11}+\N{10})}$.

@@

@@

% \begin{algorithm}[h]
% 	\caption{Class Minimization Algorithm}
% 	\label{alg:ls}
% 	\begin{flushleft}
% 		\textbf{Input:} $proj$ - project name \\
% 		\hspace{2.75em} $bug$ - bug number\\
% 		\hspace{2.75em} $k$ - number of top ranked classes\\
% 		\hspace{2.75em} $stk$ - expected stacktrace\\
% 	 \textbf{Output:} Top-k classes minimization for each test \\
% 	\end{flushleft}
% 	\begin{algorithmic}[1]
% 		\Procedure{lithium-slicer}{$proj$, $bug$, $k$, $stk$}
% 			\State $testsInfo \leftarrow$ generateInputs($proj$, $bug$, $k$, $stk$)
% 			\ForAll {$t \in testsInfo$}
% 				\State $classes \leftarrow$ getClasses($proj$, $bug$)
% 				\ForAll {$c \in classes$}
% 					\State $unc \leftarrow$ removeComments($c$)
% 					\State $min \leftarrow$ MozillaLithium($iFunc$, $unc$, $t$, $stk$)
% 				\EndFor
% 				\State $slicer \leftarrow$ getLocation($c$, $t$, $min$)
% 			\EndFor
% 			\State \Return $slicer$
% 		\EndProcedure
% 	\end{algorithmic}
%
% \end{algorithm}
%
% \begin{figure*}[t]
% 	\centering
% 	\includegraphics[width=1.0\textwidth]{figures/lithium.pdf}
% 	\caption{Simple illustration of the process to obtain the top-5 minimized classes from a project carrying a bug}
% \end{figure*}
%

The \cs{} technique is somewhat based on the \textit{statement deletion} mutant
operator of the mutation-based testing methodology. This operator was designed
based on the idea that each statement has an effect on the program output (i.e.,
oracle). Thus, testers are encouraged to design test sets that cause all
statements to be executed and generate outputs that are different from the
program under test.

@@

Considering some of the stacktrace information about the path taken to reveal
the fault solved the issue of generic oracles (e.g., NPE or generic
AssertionFailedError expections as the one presented before) mentioned in
Theorem~\ref{the:1}. There are faults whose outputs could slightly differ from
one execution to another. For example, exceptions including information of
memory addresses. For these cases, it was developed a mechanism to compare the
rest of the output ignoring the identifier of the memory address. This heuristic
also handles StackOverflow oracles.
%
\section{Evaluation}
\label{sec:eval}
%
To evaluate our approach, we compare the cost of diagnosing
a collection of faults using \sfl{} versus \comb{}.
%
\subsection{Objects of Analysis}\label{sec:analysis}

@@

		\caption{Faults of omissions' distribution}
		\label{fig:foos}
\end{figure}

@@

To evaluate the effectiveness of \comb{}, following related work, we compared its
performance with \sfl{} for each fault from the dataset, for $k\in\{5,10\}$,
where $k$ is the number of classes comprising the highest faulty statements. These
classes are obtained directly from the ranking which is calculated through \sfl{}.
For each fault, spectra is updated and the fault ranking is re-calculated but only
considering the $k$ highest faulty classes. We refer to these variations as SFL$^{k}$.
After slicing each class of the top $k$ of the highest faulty classes for each fault,
the initial spectra is updated based on the slicer output and the fault ranking is
re-calculated. We refer to these variations as \combpar{k}. We
compared \combpar{5} and \combpar{10} against \sfl{}$^{5}$ and \sfl{}$^{10}$ using
the commonly used metric $\Delta{}C = C(\textrm{SFL$^{k}$}) - C(\textrm{\combpar{k}})$,
where $C$ is the cost of diagnosis. $C$ is the mean ranking of all buggy statements.
$\Delta C >0$ means that the position of the faulty statement has decreased
in the ranking after applying the \combpar{k} approach. Hence, reducing the cost of
locating the bug. $\Delta C <0$ means that the
faulty statement may be removed by \combpar{k}. $\Delta C=0$ means that the faulty statement
remained in the same position.  We also assume a perfect bug understanding, i.e., if the developer
sees the bug in the ranking, he can say it is the faulty one.

@@

		\bottomrule
	\end{tabular}
	\caption {\ds{} reduction for $k=5$ and $k=10$}
	\label{tab:red}
\end{table}
\normalsize

@@

In \textbf{RQ1}, our concern is to evaluate how often \ds{} misses faulty statements,
since it is one of the main issues of dynamic slicing techniques. Table \ref{table:fsws}
presents the number of faults where at least one of the faulty statements appears
in the final reports of both techniques, \sfl{}$^{k}$ and \combpar{k}. We only search
for faulty statements in the top $5$ ($k=5$) and $10$ ($k=10$) of the highest faulty
classes of the projects. \combpar{k} finds the faulty statements faster than
\sfl{}$^{k}$. Using \combpar{10} we are able to find more $20$ faulty statements than
when only using \sfl{}$^{10}$. \combpar{10} only misses $11.8\%$ of the faulty statements.

>>>>

%% Although the technique has been intensively investigated in
%% research, few use cases gained notoriety---WhyLine~\cite{Ko:2008:DRA:1368088.1368130} being an
%% exception.

Model-based software diagnosis (MBSD)~\cite{REITER198757,DEKLEER200325} and Spectrum-Based Fault Localization
(\sfl{}) are two very popular techniques that leverage different principles
to automate fault isolation. An example of a MBSD is Dynamic
Slicing (DS)~\cite{Agrawal:1990:DPS:93542.93576} as well as some work~\cite{wotawa2002model,Mayer:2008:EMM:1642931.1642950,mayer2008prioritising,Perez:2018:LQR:3304889.3304927},
a technique that traces back statements in the code that influence some 
point of interest, such as the evaluation
of a failing assertion. In contrast to \ds{}, \sfl{}~\cite{7390282}
does not analyze code. Instead, it is a black-box, statistical fault
localization method that computes suspiciousness values associated
with program entities (e.g., methods) based on coverage information gathered
during the execution of test cases as well as the
output of test cases. More precisely, \sfl{} uses coverage information
of passing and failing test cases to identify likely faulty
statements and produces on output a list of program entities ranked in
decreasing order of suspiciousness.
\ds{} and \sfl{} are complementary debugging techniques~\cite{DBLP:conf/sac/AbreuGZG08}.
Intuitively, \ds{} identifies irrelevant parts of the code whereas \sfl{}
ranks the relevant parts of the code.

@@

The first question addresses the important problem of missing faulty
statements. Naturally, the consequence of that problem for debugging
is that developers would never be able to locate faults. Different
slicers can miss faulty statements for different reasons. In this
paper, we used Critical Slicing~\cite{DeMillo:1996:CSS:229000.226310},
which, in principle, could miss statements because of imprecise
oracles that guide the slicing process. We evaluated how often that
issue manifests in our experiments. The second research questions
effectively evaluates the impact of the combination. Our
results show that \Sof{@todo: sum up the results}.

The contributions of this work are the following ones:
\begin{itemize}
	\item A study, using real-world applications and bugs, on the combination
    of \ds{} and \sfl{} for bug localization of \emph{Java} faulty programs.
   \item A tool that implements \comb{}.
\end{itemize}

@@

Several types of spectra exist.  The most commonly used is called
hit-spectrum, where the coverage matrix is encoded in terms of binary
\emph{hit} (1) and \emph{not hit} (0) flags, \ie{}, $\A{ij} = 1$ if
$t_i$ covers $c_j$ and $\A{ij} = 0$ otherwise.  \sfl{} takes as input
the pair $(\mathcal{A},e)$ and produces on output a list of components
ranked by their faulty suspiciousness. To that end, the first step of
the technique consists of determining what columns of the matrix $A$
resemble the error vector $e$ the most.  For that, an intermediate
component frequency aggregator $n_{pq}(j)$ is computed $n_{pq}(j) =
|\{i\mid \A{ij}=p \wedge e_i=q\}|$. $n_{pq}(j)$ denotes the number of
runs in which the component $j$ has been active during execution ($p =
1$) or not ($p=0$), and in which the runs failed ($q = 1$) or passed
($q = 0$).  For instance, $n_{11}(j)$ counts the number of times
component $j$ has been involved ($p = 1$) in failing executions ($q =
1$), whereas $n_{10}(j)$ counts the number of times component $j$ has
been involved in passing executions. We then calculate similarity to
the error vector by means of applying \emph{fault predictors} to each
component to produce a score quantifying how likely it is to be
faulty.  Components are then ranked according to such likelihood
scores and reported to the user. Ochiai is one of those fault
predictors that has shown to perform
well~\cite{7390282,Pearson:2017:EIF:3097368.3097441}. The Ochiai
formula is given by the following equation
$\textit{ochiai}\,$=$\,\N{11}/\sqrt{(\N{11}+\N{01}) * (\N{11}+\N{10})}$.
%The Ochiai formula is given by~\cite{DBLP:conf/prdc/AbreuZG06}
%\[ochiai = \frac\]

@@

%% \vspace{3mm}
%% \begin{myframe} \label{box:1}
%% \textbf{java.lang.NullPointerException}, \\
%% at org.jfree.data.general.junit.DatasetUtilitiesTests.\\
%% testBug2849731$\_$3(DatasetUtilitiesTests.java:1299), \\
%% ...
%% \end{myframe}

@@

% \begin{algorithm}[h]
% 	\caption{Class Minimization Algorithm}
% 	\label{alg:ls}
% 	\begin{flushleft}
% 		\textbf{Input:} $proj$ - project name \\
% 		\hspace{2.75em} $bug$ - bug number\\
% 		\hspace{2.75em} $k$ - number of top ranked classes\\
% 		\hspace{2.75em} $stk$ - expected stacktrace\\
% 	 \textbf{Output:} Top-k classes minimization for each test \\
% 	\end{flushleft}
% 	\begin{algorithmic}[1]
% 		\Procedure{lithium-slicer}{$proj$, $bug$, $k$, $stk$}
% 			\State $testsInfo \leftarrow$ generateInputs($proj$, $bug$, $k$, $stk$)
% 			\ForAll {$t \in testsInfo$}
% 				\State $classes \leftarrow$ getClasses($proj$, $bug$)
% 				\ForAll {$c \in classes$}
% 					\State $unc \leftarrow$ removeComments($c$)
% 					\State $min \leftarrow$ MozillaLithium($iFunc$, $unc$, $t$, $stk$)
% 				\EndFor
% 				\State $slicer \leftarrow$ getLocation($c$, $t$, $min$)
% 			\EndFor
% 			\State \Return $slicer$
% 		\EndProcedure
% 	\end{algorithmic}
%
% \end{algorithm}
%
% \begin{figure*}[t]
% 	\centering
% 	\includegraphics[width=1.0\textwidth]{figures/lithium.pdf}
% 	\caption{Simple illustration of the process to obtain the top-5 minimized classes from a project carrying a bug}
% \end{figure*}
%
% \subsection{Function of Interest }
% \label{sec:funcofint}
%
%

The \cs{} technique is somewhat based on the \textit{statement deletion} mutant
operator (SSDL) of the mutation-based testing methodology. This operator
was designed based on the idea that each statement has an effect on the program
output (i.e., oracle). Thus, testers are encouraged to design test sets that cause
all statements to be executed and generate outputs that are different from the
program under test.

@@

Considering some of the stacktrace information about the path taken to reveal the fault 
solved the issue of generic oracles (e.g., NPE or generic AssertionFailedError expections 
as the one presented before) mentioned in Theorem~\ref{the:1}. There are faults whose outputs 
could slightly differ from one execution to another. For example, exceptions including information 
of memory addresses. For these cases, it was developed a mechanism to compare the rest of the output 
ignoring the identifier of the memory address. This heuristic also handles StackOverflow oracles.

%
% This information can be deducted from the defects stack traces which show the stack of functions called until the thrown of an exception. The interesting function compares both stack traces (from the test under evaluation and the expected one) until the call where the test fails. For each chunk under evaluation, the test is executed after the chunk being deleted from the source code . Both test ($testStk$) and expected ($expStk$) stack traces are filtered. We only consider stack traces until the line where the main exception is launched. These operations are performed by line 3 and 4 from Algorithm 2. Finally, both messages are compared and if equal the chunk will be deleted from the project, i.e., the chunk is not interesting to reveal the fault of the test.

\section{Evaluation}
\label{sec:eval}

To evaluate our approach, we compare the cost of diagnosing
a collection of faults using \sfl{} versus \comb{}.

\subsection{Objects of Analysis}\label{sec:analysis}

@@

		\caption{FAULT$\_$OF$\_$OMISSION distribution in the \dfj{} dataset}
		\label{fig:foos}
\end{figure}

@@

To evaluate the effectiveness of \comb{}, following related work, we compared its
performance with \sfl{} for each fault from the dataset, for $k\in\{5,10\}$, 
where $k$ is the number of classes comprising the highest faulty statements. These 
classes are obtained directly from the ranking which is calculated through \sfl{}. 
For each fault, spectra is updated and the fault ranking is re-calculated but only 
considering the $k$ highest faulty classes. We refer to these variations as SFL$^{k}$. 
After slicing each class of the top $k$ of the highest faulty classes for each fault, 
the initial spectra is updated based on the slicer output and the fault ranking is 
re-calculated. We refer to these variations as \combpar{k}. We
compared \combpar{5} and \combpar{10} against \sfl{}$^{5}$ and \sfl{}$^{10}$ using the
following metric:

\begin{equation}
    \Delta{}C = C(\textrm{SFL$^{k}$}) - C(\textrm{\combpar{k}})
\end{equation}

\noindent
where $C$ is the cost of diagnosis. $C$ is the mean ranking of all buggy statements.
$\Delta C >0$ means that the position of the faulty statement has decreased
in the ranking after applying the \combpar{k} approach. Hence, reducing the cost of
locating the bug. $\Delta C <0$ means that the 
faulty statement may be removed by \combpar{k}. $\Delta C=0$ means that the faulty statement 
remained in the same position.  We also assume a perfect bug understanding, i.e., if the developer
sees the bug in the ranking, he can say it is the faulty one.

@@

		\bottomrule
	\end{tabular}
	\caption {\ds{} reduction performance when considering $k=5$ and $k=10$}
	\label{tab:red}
\end{table}
\normalsize

@@

In \textbf{RQ1}, our concern is to evaluate how often \ds{} misses faulty statements, 
since it is one of the main issues of dynamic slicing techniques. Table \ref{table:fsws} 
presents the number of faults where at least one of the faulty statements appears 
in the final reports of both techniques, \sfl{}$^{k}$ and \combpar{k}. We only search
for faulty statements in the top $5$ ($k=5$) and $10$ ($k=10$) of the highest faulty 
classes of the projects. \combpar{k} finds the faulty statements faster than 
\sfl{}$^{k}$. Using \combpar{10} we are able to find more $20$ faulty statements than
when only using \sfl{}$^{10}$. \combpar{10} only misses $11.8\%$ of the faulty statements
whereas \sfl{}$^{10}$ misses $17.9\%$ of the faulty statements. Thus, an improvement
of $5.2\%$.
